---
title: "Clémence_Huang"
author: "Clémence HUANG"
format: html
editor: visual
---

## Types and classes

### Types
```{r}
typeof(1.5)
typeof("toto")
typeof('also between single quotes')
typeof(TRUE)
typeof(1L)
typeof(1)
typeof(typeof)
typeof(list())
```

### Classes
```{r}
iris
typeof(iris)
class(iris)
```

## Vectors
Many things in R are vectors( and most of the other things are lists).
A vector is a collection of values of the same type (possibly a single value).
```{r}
x <- 15.25
length(x)
```

The `c` function can be seen as a creation function or, more accurately, as a vector concatenation function.

```{r}
y <- 15.25
c(x,y)
```

### Character vectors
```{r}
z <- "Bob"
length(z)
```

Salvation comes from the `stringr` package.
```{r}
library(stringr)
```

```{r}
str_length(z)
```

```{r}
complicated <-"💕😁😉😉😆R是一门很好的编程语言"
length(complicated)
str_length(complicated)
```

```{r}
some_values <- c("💕😁😉😉😆","R是一门很好的编程语言")
some_values
length(some_values)
str_length(some_values)
str_to_upper(some_values)
```

### Indexing
```{r}
x <- c(6,5,1,2,3,4)
y <- c(6,5,1,2,3,4)
x[3]
```

```{r}
x[4] <- 15
x
```

Assignment can be understood as a value semantic assignment : `x` and `y` are different vectors with the same initial content.

```{r}
y <- x
```

Therefore, modifying `x` has no effect on `y`. 

Indexing is made with vectors of any length.
```{r}
x
x[c(2,2,1,4,3)]
```

```{r}
x[0]
x[7]
x[10]
```

```{r}
typeof(NA_integer_)
NA_integer_
```

```{r}
x
x[c(4,1,0,15)]
```

```{r}
c(numeric(0),5,numeric(0))
```

```{r}
y <- c("A","B","C")
y
y[c(TRUE,FALSE,TRUE)]
```

```{r}
x
x[c(1,4)] <- c(-34,22)
x
x[c(2,10)] <- c(123,456)
x
```





