---
title: "Valentine LLR"
author: "Valentine Petzold"
format: html
---

## Types and classes 
### Types
```{r}
typeof(1.5)#double: numerical value
typeof("toto")
typeof('also between single quotes')
typeof(TRUE)
typeof(1L)
typeof(1)
typeof(typeof)
typeof(list())
```
types defined set with associated operations.
double roughly real numbers.
everything id a double if you don't tell R integer.
logical for testing.
### Classes
```{r}
iris
typeof(iris)
class(iris)
class(1)#we don't really care just concept to have in mind
```
## Vectors
### Introduction
Many things in R are vectors (and most of the other things are lists).
Vector: a collection of values of the same type.

```{r}
x<-15.25
length(x)# all vectors have a length
x[1] # access to elements
```
```{r}
length(length(x))
typeof(length(x))

```
```{r}
y<-c(1,-2,45,1.5,-1e-10)
y
typeof(y)
y[2]
length(y)
```
```{r}
c(x,y) #combine vectors
```

The 'c' function can be seen as a creation function or, more accurately, as a vector concatenantion function.*

### Character vectors

This is not R strength...

```{r}
z<- "Alice"
length(z)
```
In R everything is a vector so here only 1 not 5.
pearl language

Salvation comes from the 'stringr' package.

```{r}
library(stringr)
```

```{r}
str_length(z)
```
```{r}
complicated<-"ðŸ’•ðŸ˜ðŸ˜‰ðŸ˜‰ðŸ˜†"
complicated
length(complicated)
str_length(complicated)
```

```{r}
#some_values<-c("")
#some_values
#length(some_values)
#str_length(some_values)
```

### Indexing

```{r}
x<-c(6,5,1,2,3,4)
x[3]
```
```{r}
x[4]<-15#change the forth element
x
```
Assignment can be understood has a value semantic assignment: 'x' and 'y' are different vectors with the same initial content.
```{r}
y<-x
y
```
Therefore, modifying 'x' has no effect on 'y'.
```{r}
x[1]<-18
x
y
```
Indexing is made with vectors of any length.
```{r}
x
x[c(2,2,1,4,3)]
```
```{r}
x[0]#only one that does not work bc negative numbers work as well, no possible meaning -> empty numeric vector
x[7]
x[10]#too large gives NA, no value there -> not available
```
```{r}
typeof(numeric(0))
length(numeric)
```
```{r}
typeof(NA)
length(NA)
```
```{r}
typeof(NA_integer_)
NA_integer_
```
```{r}
x
x[c(4, 1, 0, 15)]
```
```{r}
c(numeric(0), 5, numeric(0))
```
```{r}
x
x[c(-1,-4,-12)]#remove
```

```{r}
y<-c("A","B","C")
y
y[c()]
```

